\protect Connecting every bit of knowledge: \\ The Structure of Wikipedia's first link network

### 2016-02-28 writeup/wikipedia-network.tex
abstract wording tweaks

\begin{abstract}
  \protect
Apples, porcupines, and the most obscure Dylan song---is every topic a few clicks from Philosophy? 
Within Wikipedia, the surprising answer is yes: nearly all 
paths lead to Philosophy.
Wikipedia is the largest, most meticulously indexed collection of human knowledge ever amassed. 
More than information about a topic, Wikipedia is a web of naturally emerging relationships.  
By following the first link in an article, we connect entries to form a directed network: Wikipedia's First Link Network. 
Here, we study the English edition of Wikipedia's First Link Network for insight into how the many topics, ideas, people, objects, and events
are related and organized.  
We algorithmically parse all 4.7 million articles to construct a map of Wikipedia's First Link Network. 
We traverse every possible path through the network, 
measuring the accumulation of first links, path lengths, basins, cycles, and the influence a particular article exerts in shaping the 
network.
We discover many scale-free distributions describing path length, 
accumulation, and influence; find Philosophy at a salient center; and uncover a flow from specific to general 
culminating around fundamental notions such as Community, State, and Science. 
Curiously, we also observe a gravitation towards topical articles including Health Care and Fossil Fuel.
These findings enrich our view of the connections and structure of
Wikipedia's ever growing store of knowledge.

@@

The authors of a Wikipedia article choose where and whether to include a 
reference to another Wikipedia article
in the HTML markup.
For example, the authors of 
\href{https://en.wikipedia.org/wiki/Train}{``Train''}
chose
``Amtrak's Acela Express,'' ``steam,'' and ``head-end power'' 
among others as relevant articles to reference in describing ``Train''
\cite{wiki_train}
.
By focusing our attention on the main body of an article---excluding
elements in the side bars and headings---we capture the introduction to the topic.
The first link marks the earliest moment in a topic's introduction where the 
authors choose to directly reference another article. 
While the many links in an article may reference relevant details to a topic, 
the first link is an association within the introductory 
description. While 
``Amtrak's Acela Express,'' ``steam,'' and ``head-end power'' 
are links detailing particulars, the first link, ``rail transport,''
is the topic the authors associate with ``Train'' in the introduction. 
Collectively, the first links provide a pragmatic and interpretable 
means to connect each article to another.

By following the first hyperlink to another article in the English edition of
Wikipedia as of November, 2014, we connect the ideas in one article to those in another, ultimately forming a directed network: 
{\it Wikipedia's First Link Network} (FLN).

\begin{figure*}[tp!]
  \includegraphics[width=\textwidth]{graphics/Train.pdf}  
  \caption{
    \textbf{First Link Path For ``Train.''}
    We follow the First Link to another Wikipedia article
    in the main body of the article---the area
    inside the green rectangle, which excludes 
    side bar elements, the navigation bar and title. 
    In this example, the First Link to another Wikipedia article is ``Rail Transport.'' 
    We can again select the First Link on the ``Rail Transport''
    article, repeating the process to 
    form a path of First Links.
  }
  \label{fig:Train First Links}
\end{figure*}

What information could the First Links possibly reveal?
Inspired by the claim that the majority of First Links lead to the 
``Philosophy'' article---popularized by an 
\href{https://xkcd.com/903/}{xkcd}
comic and subsequently
discussed in blog posts 
\cite{mat_blog, Ilmari_first_links, xkcd}
---we holistically study 
Wikipedia's FLN as a map relating areas of human knowledge. 
The map is a hierarchical taxonomy where, for example, ``Train'' links to a parent node, 
``Rail Transport,'' while many child nodes, ``Steel'' and ``Horsepower,'' link to 
``Train.'' Unlike previous taxonomies
created by individuals
\cite{locke, descartes, aristotle}
or a select group of individuals 
\cite{hist_thesaurus}, 
the organization of ideas in the FLN 
does not emerge from a centralized effort. 
Rather each link in the FLN emerges from the choice of each author and ensuing edits.
Collectively, the FLN is a wealth of relations among inventions, places,
figures, objects, and events across space and time.

Our goal is to study the structure of the FLN for insight into how the information on Wikipedia is organized and related.
We develop metrics to capture 
the dominant features of the FLN's structure.
We measure dynamics of the FLN as a flow, quantifying 
aspects of the FLN from the accumulation of First Links around articles 
to the influence an article exerts in shaping the FLN.
Together with cycles, in-degree, depth, and the content of the articles, 
we build our analysis of the relations among the ideas in Wikipedia.

@@

Our greater interest is lies in the richer dynamics of how many articles fit together---how links create a flow through the FLN.
We construct a path through the FLN with each article 
leading to the next through the first link. 
For example ``Train'' has a first link to ``Conveyance of Passengers and Goods,'' which is itself
an article with a first link to ``Goods'' and so on. 
As shown in Fig.~\ref{fig:Train First Links}, the path starting at ``Train'' 
contains ``Goods,'' ``Economics,'' ``Social Science,''
leading ultimately to ``Philosophy''.

@@

\begin{figure}[tp!]
  \includegraphics[width=\columnwidth]{graphics/traversals_per_article.png} 
  \caption{
    \textbf{Distribution of Traversal Visits.}
We fit the distribution of traversal visits against article rank with a log-log (base 10) linear regression model log-log model and uncover two regimes: 
with a cutoff rank of $10^{5}$.
The top regime ($\log_{10}$(rank) $< 5$) has a Pearson's Correlation 
Coefficient of $-0.99$ with a corresponding power law rank exponent 
$\alpha = -1.23$ and a size exponent $\gamma = 0.187$.
The horizontal flattening around the highest
ranking articles is a result of the cyclic structure (see discussion on cycles).}
  \label{fig:Distribution of Visits}

@@

In Fig.~\ref{fig:Distribution of Visits}, we show a log-log (base 10) plot for the entire dataset: traversal visits against rank. 
We fit the data to a two regime log-log linear regression with a cutoff rank 
of $10^{5}$.
The top regime ($\log_{10}$(rank) $< 5$) has a Pearson's Correlation 
Coefficient of $-0.99$ with a corresponding power law rank exponent 
$\alpha = -1.23$ and a size exponent $\gamma = 0.187$.
On the aggregate, the distribution suggests
a handful of the highest ranking articles contain a disproportionate number of traversal visits, while most have none. The skew in the distribution is not terribly surprising when considering the heuristic of how the links flow: from specific to general.

@@

@@

\begin{figure*}[tp!]
  \includegraphics[width=\textwidth]{graphics/2_cycles.pdf}
  \caption{
    \textbf{Highest ranking 2-cycles.}
We identify pairs of articles whose first links point to one another, forming
a 2-cycle. We then rank each pair of articles by the total number of 
traversal visits to gauge the most referenced groups of two articles linked
to each other. The area of the circles in the figure indicates the number of traversal visits. We find 2-cycles often capture synonyms or articles representing nearly the 
same concepts as opposed to distinct concepts.}
  \label{fig:2-cycles}
\end{figure*}
We first identified 2-cycles, meaning a pair of articles with first link pointing to one another.
Of the 11 million articles, 84,000 are 2-cycles. 
The highest ranking 2-cycles by traversal visits tend to be synonyms (or nearly so) rather than distinct, yet connected topics:
``Health Care'' and ``Medical Case Management,'' ``Broadcasting House,'' and ``BBC,'' ``Secondary Education'' and ``Secondary School'' 
(see Fig.~\ref{fig:2-cycles}).

@@

\begin{figure}[tp!]
  \includegraphics[width=\columnwidth]{graphics/top_funnels.png}
  \caption{
    \textbf{Funnels.}
We represent the highest-ranking articles by the number of traversal 
funnels to gauge the influence each article exerts in shaping the 
structure of the FLN. 
The area of each circle is sized in proportion to the number of 
traversal funnels each article has. 
We find ``Philosophy'' exerts an overwhelming proportion
of the influence, with other abstract notions and topical concepts ranking
next.}
  \label{fig:Funnels}
\end{figure}
To analyze the influence an article exerts in shaping the 
structure of the FLN, we compute the number of traversal funnels for each.
Articles directing more paths exert a greater influence over the structure
of the FLN by increasing the accumulation of first links
on a particular path. By measuring traversal funnels, we distinguish between an article that simply happened to fall within a cycle from an article funneling 
many first links.

\begin{figure}[tp!]
  \includegraphics[width=\columnwidth]{graphics/funnels_distribution.png}
  \caption{
    \textbf{Distribution of Traversal Funnels.}
We fit the distribution with a linear log-log model by considering the log (base 10) transformed rank of each article against log (base 10) transformed the number of traversal funnels. 
We find two regimes with the top regime (log(rank) < 4) 
well-explained by a linear fit, yielding Pearson's 
Correlation Coefficient of $-0.99$ and a 
corresponding power law rank exponent $\alpha = -1.08$
and a size exponent $\gamma = 0.074$. The top regime 
explains the distribution of traversal funnels for the $17, 821$ 
articles with at least one traversal funnel. The bottom 
horizontal regime corresponds to the more than $99\%$ of articles
which hold zero traversal funnels.}
  \label{fig:Funnels Distribution}
\end{figure}
Ranking articles by the number of traversal funnels we find 
``Philosophy'' to be by far the highest-ranking article with 
$7.37$ million paths
(see Fig.~\ref{fig:Funnels}).
Of any article, the number of traversal funnels Philosophy holds exceeds 
all others by more than two orders of magnitude.
The ``Philosophy'' cycle which contains ``Existence,'' ``Awareness,'' ``Reality,'' 
and similar articles accumulates the overwhelming proportion of its 
references through ``Philosophy'': $7.37$ million of the $7.4$ million references
are funneled through ``Philosophy''.
Second on the list of highest-ranking articles by traversal funnels is 
``Presentation'' with only $30$ thousand paths. Similarly abstract 
concepts also rank highly such as ``Tree of life'' (30 thousand), 
``Reality'' (13 thousand), and ``Jurisdiction'' (3 thousand).

@@

\end{figure}
Wikipedia released an API to measure article popularity by page views
starting November 2015. Page views add another dimension to our
findings by revealing 
how users view the information on Wikipedia. 
We measure popularity as the total number 
of page views in the English edition of Wikipedia in the month of 
October 2015---the earliest full month for which the data is available. 
We find the highest ranking 1000 articles (by traversal visits) have an average of
$70$ thousand page views in October with high variation: the standard deviation 
is $1.1$ million page views. 
The number of page views has a skewed distribution with $75\%$ of articles
reaching fewer than $73$ thousand page views in October.
The article for ``United States'' has the most page 
views of the highest ranking $1000$ articles by traversal visits with 
$1.4$ million views in October. The next most popular articles are 
``India,'' ``World War II,'' and the ``United Kingdom'' each with $0.7-1.03$ million page views.

@@

Among our observations is the multiple appearance of scale-free distributions within the network. 
Few articles have most traversal visits, few paths have an exceptionally long path length, and even fewer
articles are responsible for funneling most paths. When measured against the traversal funnels, 
``Philosophy'' emerges as an exceptional article by orders of magnitude. 
Nevertheless, many other foundational concepts emerged naturally within the First Link Network. 
Basins around ``Community'', ``State'', and ``Science'' reveal a foundational structure within the network. 
More curious is the emergence of recently prominent political and economics topics such as ``Fossil Fuel'' and ``Health Care'' 
within the highest ranking funnels. 
Wikipedia seems to reflect not only timeless foundations, but also the topical (at least within English speaking society).

Future work could examine other language versions of Wikipedia for potentially telling cultural or regional differences as well as expand the network to more than the First Link.
These findings also form the basis for the creation of a taxonomy where 
every idea, event, or object sits within a hierarchy of connected notions.
The taxonomy would extend a traditional word thesaurus beyond mere synonyms to a related hierarchy of concepts.
Applications could range from an enhanced network of ideas to psychological insights into how humans form associations.
Specifically, an ever-evolving reference of related hierarchical concepts could be applied to search engine algorithms 
or natural language processing.

@@

Wikipedia renders and stores articles in MediaWiki markup, a markup language with syntax and keywords to format and mark elements in a page. Along with special syntax for links, MediaWiki markup includes templates for audio files, images, and side-bar
information.
While a human could accurately identify the First Link, to map the entire First Link Network of 11 million articles, we needed to programmatically untangled the body text from side-bar, header box, and invalid link elements.

While some libraries exist for MediaWiki Markup,
Approaches using existing libraries led to several bugs 
including trouble with nested links, nested parenthesis, unclosed tags, escape characters 
as well as compatibility with other libraries used to parse the XML.
Consequently,  we developed an algorithm for parsing the First Link in the XML version of each article.
Our parsing algorithm aimed to: 
1) accurately identify the First Link among other page elements, and 
2) efficiently do so---that is without needing for several passes through the data.
To process an article in a single pass, we developed a hierarchical system of flags:
\begin{figure}[tp!]
  \includegraphics[width=\columnwidth]{graphics/flags.pdf}  
  \caption{
    \textbf{Parsing Algorithm of Wikipedia's XML dump.}
The highest flag in the hierarchy indicates a Wikimedia template used to mark an element in the side bar, display an image, link to an external file, or another Wikimedia project outside of Wikipeida. Next, to catch any remaining elements outside the main body we have a second flag for <ref>, <div> elements. Finally, we identify parenthesis to ensure we do not capture a link to a pronunciation key.}
  \label{fig:parsing algorithm}
\end{figure}

The algorithm loops in three-character chunks to account for potentially nested elements, 
shifting by one character steps through the article markup.
If any markup triggers for a flag are detected, a flag is raised. 
Once a flag is raised, we stop processing and proceed to the next character
until the flag's closing markup.
A First Link is identified only if Flags 1, 2, and 3 are all off.
In this case, the entire link is retrieved. 
We then confirm the link is valid by filtering for MediaWiki keywords indicating external page or other projects
as well as common file extensions for 
images, audio files, and the like 
\cite{media_wiki_templates}.
The First Link of an article is then the earliest valid link with unraised flags.

To process the entirety of Wikipedia, we distributed the parsing and processing of the XML dump
across 112 cores of the UVM supercomputer cluster
\cite{vacc}
We then joined the results to form a hash table containing every Wikipedia article and its corresponding
First Link. The resulting network map is the basis of our analysis.

>>>>

\begin{abstract}
  \protect
Apples, porcupines, and the most obscure Dylan song---is every topic a few clicks from Philosophy? 
Within Wikipedia, the surprising answer is yes: nearly all 
paths lead to Philosophy.
Wikipedia is the largest, most meticulously indexed collection of human knowledge ever amassed. 
More than information about a topic, Wikipedia is a web of naturally emerging relationships.  
By following the first link in an article, we connect entries to form a directed network: Wikipedia's First Link Network. 
Here, we study the English edition of Wikipedia's First Link Network for insight into how the many 
inventions, places, people, objects, and events
are related and organized.  
We algorithmically parse all 4.7 million articles to construct a map of Wikipedia's First Link Network. 
We traverse every possible path through the network, 
measuring the accumulation of first links, path lengths, basins, cycles, and the influence a particular article exerts in shaping the 
network.
We discover scale-free distributions describe path length, 
accumulation, and influence. Far from dispersed, first links disproportionately accumulate 
at a few articles---flowing from specific to general and culminating around fundamental notions such as
Community, State, and Science. Even fewer articles most influence the network's structure with Philosophy 
shaping more paths than any other article by two orders of magnitude. 
Curiously, we also observe a gravitation towards topical articles such as 
Health Care and Fossil Fuel. 
These findings enrich our view of the connections and structure of
Wikipedia's ever growing store of knowledge.

@@

The authors of a Wikipedia article choose where and whether to include a 
reference to another Wikipedia article
in the HTML markup.
For example, the authors of 
\href{https://en.wikipedia.org/wiki/Train}{``Train''}
chose
``Amtrak's Acela Express,'' ``steam,'' and ``head-end power'' 
among others as relevant articles to reference in describing ``Train''
\cite{wiki_train}
.

By focusing our attention on the main body of an article---excluding
elements in the side bars and headings---we 
systematically capture the core description of a topic.
Within the body text, the first link marks the earliest moment in a topic's introduction where the 
authors choose to directly reference another article. 
While many links reference relevant details, 
the first link is an association within a topic's initial 
description. While 
``Amtrak's Acela Express,'' ``steam,'' and ``head-end power'' 
are links detailing particulars, the first link, ``rail transport,''
is the topic the authors associate with ``Train'' in the introduction. 
``Banana'' has a first link to ``fruit,'' ``Bob Dylan'' has a first link to 
``Blowing in the Wind,'' and ``Physics'' has a first link to ``natural science.''
Collectively, the first links provide a pragmatic and interpretable 
means to connect each article to another.

By following the first hyperlink to another article in the English edition of
Wikipedia as of November, 2014, we connect the topic of one article to that of another, ultimately forming a directed network: 
{\it Wikipedia's First Link Network} (FLN).

\begin{figure*}[tp!]
  \includegraphics[width=\textwidth]{graphics/Train.pdf}  
  \caption{
    \textbf{First Link Path For ``Train.''}
    We follow the first link to another Wikipedia article
    in the main body of the article---the area
    inside the green rectangle, which excludes 
    side bar elements, the navigation bar and title. 
    In this example, the first link to another Wikipedia article is ``Rail Transport.'' 
    We can again select the first link on the ``Rail Transport''
    article, repeating the process to 
    form a path of first links.
  }
  \label{fig:Train First Links}
\end{figure*}

The FLN is a wealth of relations among inventions, places,
figures, objects, and events across space and time.
``Train'' for example, links to a parent node, ``Rail Transport,''
while many child nodes such as ``Steel'' and ``Horsepower'' link to ``Train.''
Unlike previous taxonomies created by individuals
\cite{locke, descartes, aristotle, hist_thesaurus},
the relations in the FLN emerge without a centralized effort 
as the aggregate of each article's authors choice of first link.

Our goal is to study the structure of the FLN for insight into how the information on Wikipedia is organized and related.
We develop metrics to capture 
the dominant features of the FLN's structure.
We measure dynamics of the FLN as a flow, quantifying 
aspects of the FLN from the accumulation of first links around articles 
to the influence an article exerts in shaping the FLN.
Together with cycles, in-degree, depth, and the content of the articles, 
we build our analysis of the relations among the ideas in Wikipedia.

@@

Our greater interest is lies in the richer dynamics of how many articles fit together.
Inspired by the claim that the majority of first links lead to
``Philosophy''---popularized by an 
\href{https://xkcd.com/903/}{xkcd}
comic and subsequently
discussed in blog posts 
\cite{mat_blog, Ilmari_first_links, xkcd}
---
we holistically study how the FLN  yields a flow of connections.
We construct a path through the FLN with each article 
leading to the next through the first link. 
For example ``Train'' has a first link to 
``Rail Transport,''
 which is itself
an article with a first link to 
``Conveyance of Passengers and Goods,''
and so on.
As shown in Fig.~\ref{fig:Train First Links}, the path starting at ``Train'' 
contains ``Goods,'' ``Economics,'' ``Social Science,''
leading ultimately to ``Philosophy''.

@@

\begin{figure}[tp!]
  \includegraphics[width=\columnwidth]{graphics/traversals_per_article.png} 
  \caption{
    \textbf{Distribution of Traversal Visits.}
We fit the distribution of traversal visits against article rank with a log-log (base 10) linear regression model log-log model and uncover two regimes: 
with a cutoff rank of $10^{5}$.
The top regime ($\log_{10}$(rank) $< 5$) has a Pearson's Correlation 
Coefficient of $-0.99$ 
with a corresponding power law rank exponent 
$\alpha = -1.23$ and a size exponent $\gamma = 0.187$.
The horizontal flattening around the highest
ranking articles is a result of the cyclic structure (see discussion on cycles).
The bottom regime has a Pearson's Correlation Coefficient of $-0.93$
with a corresponding power law rank exponent 
$\alpha = -0.579$ and a size exponent $\gamma = -0.727$.}
  \label{fig:Distribution of Visits}

@@

In Fig.~\ref{fig:Distribution of Visits}, we show a log-log (base 10) plot for the entire dataset: traversal visits against rank. 
We obtain an excellent fit with regimes corresponding to power law rank exponents of 
$\alpha = -1.23 (\gamma = 0.187$) and $\alpha = -0.579 (\gamma = -0.727)$.
On the aggregate, the distribution suggests
a handful of the highest ranking articles contain a disproportionate number of traversal visits, while most have none. The skew in the distribution is not terribly surprising when considering the heuristic of how the links flow: from specific to general.

@@

The highest ranking articles by traversal visits are broad topics spanning academic disciplines 
or notions fundamental to society~(Fig. \ref{fig:highest visits}): ``Science,'' ``Mathematics,'' `` Geography,'' and ``Philosophy''
as well as ``Community,'' ``State,'' ``Earth,'' ``Information,'' ``Power,'' and ``Communication.''
The first links flow towards broader topics, accumulating at of these foundational notions.
While ``Banana'' is a concrete fruit for example, 
the first links flow from ``Fruit'' to ``Botany'' to ``Biology,'' and ultimately 
culminate with ``Science'' and ``Philosophy.'' 
The first links anchor the wealth of specific knowledge on Wikipedia to 
a few notions foundational to society.

@@

\begin{figure*}[tp!]
  \includegraphics[width=\textwidth]{graphics/2_cycles.pdf}
  \caption{
    \textbf{Highest ranking 2-cycles.}
We identify pairs of articles whose first links point to one another, forming
a 2-cycle. We then rank each pair of articles by the total number of 
traversal visits to gauge the most referenced groups of two articles linked
to each other. The area of the circles in the figure indicates the number of traversal visits. We find 2-cycles often capture synonyms or articles representing nearly the 
same concepts as opposed to distinct concepts.}
  \label{fig:2-cycles}
\end{figure*}
We identify 2-cycles, meaning a pair of articles with first link pointing to one another.
Of the 11 million articles, 84,000 are 2-cycles. 
The highest ranking 2-cycles by traversal visits tend to be synonyms (or nearly so) rather than distinct, yet connected topics:
``Health Care'' and ``Medical Case Management,'' ``Broadcasting House,'' and ``BBC,'' ``Secondary Education'' and ``Secondary School'' 
(see Fig.~\ref{fig:2-cycles}).

@@

\begin{figure}[tp!]
  \includegraphics[width=\columnwidth]{graphics/top_funnels.png}
  \caption{
    \textbf{Funnels.}
We represent the highest-ranking articles by the number of traversal 
funnels to gauge the influence each article exerts in shaping the 
structure of the FLN. 
We find ``Philosophy'' exerts an overwhelming proportion
of the influence, with other abstract notions and topical concepts ranking
next.}
  \label{fig:Funnels}
\end{figure}
To analyze the influence an article exerts in shaping the 
structure of the FLN, we compute the number of traversal funnels for each.
Articles directing more paths exert a greater influence over the structure
of the FLN by increasing the accumulation of first links
on a particular path. By measuring traversal funnels, we distinguish between an article that simply happened to fall within a cycle from an article funneling 
many first links.

\begin{figure}[tp!]
  \includegraphics[width=\columnwidth]{graphics/funnels_distribution.png}
  \caption{
    \textbf{Distribution of Traversal Funnels.}
We fit the distribution with a linear log-log model by considering the log (base 10) transformed rank of each article against log (base 10) transformed number of traversal funnels. 
We find two regimes with the top regime (log(rank) < 4) 
well-explained by a linear fit, yielding Pearson's 
Correlation Coefficient of $-0.99$ and a 
corresponding power law rank exponent $\alpha = -1.08$
and a size exponent $\gamma = 0.074$. 
The bottom horizontal regime corresponds to the more than $99\%$ of articles
which hold zero traversal funnels.}
  \label{fig:Funnels Distribution}
\end{figure}
Ranking articles by the number of traversal funnels we find 
``Philosophy'' to be by far the highest-ranking article with 
$7.37$ million paths
(see Fig.~\ref{fig:Funnels}).
Of any article, the number of traversal funnels Philosophy holds exceeds 
all others by more than two orders of magnitude.
The ``Philosophy'' cycle which contains ``Existence,'' ``Awareness,'' ``Reality,'' 
and similar articles accumulates the overwhelming proportion of its 
references through ``Philosophy'': $7.37$ million of the $7.4$ million references
are funneled through ``Philosophy''.
Second on the list of highest-ranking articles by traversal funnels is 
``Presentation'' with only $30$ thousand paths. Similarly abstract 
concepts also rank highly such as ``Tree of life'' (30 thousand), 
``Reality'' (13 thousand), and ``Jurisdiction'' (3 thousand).

@@

\end{figure}
Wikipedia released an API to measure article popularity by page views
starting November 2015. 
Page views add another dimension to our
findings by constrasting
the number of users who access a particular article against
the number of traversal visits and funnels.
We measure popularity as the total number 
of page views in the English edition of Wikipedia in the month of 
October 2015---the earliest full month for which the data is available. 
We find the highest ranking 1000 articles (by traversal visits) have an average of
$70$ thousand page views in October with high variation: the standard deviation 
is $1.1$ million page views. 
The number of page views has a skewed distribution with $75\%$ of articles
reaching fewer than $73$ thousand page views in October.
The article for ``United States'' has the most page 
views of the highest ranking $1000$ articles by traversal visits with 
$1.4$ million views in October. The next most popular articles are 
``India,'' ``World War II,'' and the ``United Kingdom'' each with $0.7-1.03$ million page views.

@@

Among our observations is the multiple appearance of scale-free distributions within the network. 
Few articles have most traversal visits, few paths have an exceptionally long path length, and even fewer
articles are responsible for funneling most paths. When measured against the traversal funnels, 
``Philosophy'' emerges as an exceptional article by orders of magnitude. 
Nevertheless, many other foundational concepts emerged naturally within FLN. 
Basins around ``Community'', ``State'', and ``Science'' reveal a foundational structure within the network. 
More curious is the emergence of recently prominent political and economics topics such as ``Fossil Fuel'' and ``Health Care'' 
within the highest ranking funnels. 
Wikipedia seems to reflect not only timeless foundations, but also the topical (at least within English speaking society).

Future work could examine other language versions of Wikipedia for potentially telling cultural or regional differences as well as expand the network to more than the first link.
These findings also form the basis for the creation of a taxonomy where 
every idea, event, or object sits within a hierarchy of connected notions.
The taxonomy would extend a traditional word thesaurus beyond mere synonyms to a related hierarchy of concepts.
Applications could range from an enhanced network of ideas to psychological insights into how humans form associations.
Specifically, an ever-evolving reference of related hierarchical concepts could be applied to search engine algorithms 
or natural language processing.

@@

Wikipedia renders and stores articles in MediaWiki markup, a markup language with syntax and keywords to format and mark elements in a page. Along with special syntax for links, MediaWiki markup includes templates for audio files, images, and side-bar
information.
While a human could accurately identify the first link, to map the entire First Link network of 11 million articles, we needed to programmatically untangled the body text from side-bar, header box, and invalid link elements.

While some libraries exist for MediaWiki Markup,
Approaches using existing libraries led to several bugs 
including trouble with nested links, nested parenthesis, unclosed tags, escape characters 
as well as compatibility with other libraries used to parse the XML.
Consequently,  we developed an algorithm for parsing the first link in the XML version of each article.
Our parsing algorithm aimed to: 
1) accurately identify the first link among other page elements, and 
2) efficiently do so---that is without needing for several passes through the data.
To process an article in a single pass, we developed a hierarchical system of flags:
\begin{figure}[tp!]
  \includegraphics[width=\columnwidth]{graphics/flags.pdf}  
  \caption{
    \textbf{Parsing Algorithm of Wikipedia's XML dump.}
The highest flag in the hierarchy indicates a Wikimedia template used to mark an element in the side bar, display an image, link to an external file, or another Wikimedia project outside of Wikipeida. Next, to catch any remaining elements outside the main body we have a second flag for <ref>, <div> elements. Finally, we identify parenthesis to ensure we do not capture a link to a pronunciation key.}
  \label{fig:parsing algorithm}
\end{figure}

The algorithm loops in three-character chunks to account for potentially nested elements, 
shifting by one character steps through the article markup.
If any markup triggers for a flag are detected, a flag is raised. 
Once a flag is raised, we stop processing and proceed to the next character
until the flag's closing markup.
A first link is identified only if Flags 1, 2, and 3 are all off.
In this case, the entire link is retrieved. 
We then confirm the link is valid by filtering for MediaWiki keywords indicating external page or other projects
as well as common file extensions for 
images, audio files, and the like 
\cite{media_wiki_templates}.
The first link of an article is then the earliest valid link with unraised flags.

To process the entirety of Wikipedia, we distributed the parsing and processing of the XML dump
across 112 cores of the UVM supercomputer cluster
\cite{vacc}
We then joined the results to form a hash table containing every Wikipedia article and its corresponding
first link. The resulting network map is the basis of our analysis.
